A comparative analysis of penalized linear mixed models in structured genetic data

### 2021-02-15 main.tex
Minor tweaks

\anna{\textbf{More discussion of nonlinear models. In particular, is logistic regression practical?}} \anna{In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait. However, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward. LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation. This procedure does not generalize beyond the linear regression case where the outcome is continuous. Because of this, distinct inferential algorithms are required to extend the LMM modeling methodology to nonlinear models such as logistic regression. \citet{mandt2017sparse} generalize the LMM modeling paradigm to binary phenotypes in the interest of classification. They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference. However, to the best of our knowledge, the joint problem of sparse estimation in the presence of dependent observations has been limited to the linear regression case \citep{seeger2011large, vattikuti2014applying, rakitsch2013lasso}.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods} \label{sec:methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
\subsection{\anna{lasso}}
\anna{\textbf{More background on lasso: 'I acknowledge that this article cannot and should not review Lasso and other penalizations. But still, it is worthwhile briefly discussing what the key methodological components can do and providing references'}}
We begin by briefly reviewing the key methodological components of the lasso. In the usual linear regression set-up where there is a continuous response $\by \in \real^n$, an $n \times p$ design matrix $\bX$, and a parameter vector $\bbeta \in \real^p$, the ordinary lasso minimizes the objective function

\begin{equation}
    Q(\bbeta | \lam) = \frac{1}{2n}|| \by -  \beta_0 - \bX \bbeta ||_2^2 + \lam ||\bbeta||_1,
\end{equation}
where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm. The the first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data. The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values. The tuning parameter, $\lambda$, controls the tradeoff between these two components. For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0. In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of ordinary least squares (OLS). \anna{\textbf{should be clearer descriptions on when the reviewed approaches can/cannot be applied; be more explicit here?}} The lasso assumes observations are independent. However, the $\ell_1$ penalty of the lasso can also be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.

@@

%------------------------------%
\subsection{\anna{Genomewide case study}}
%------------------------------%

@@

\anna{\textbf{penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.} The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods. In a real data scenario, the true number of causal SNPs would not be known a priori. As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion. As of yet there is no clear consensus on the optimal way to select a penalty parameter for penalized LMMs.}

@@

\anna{\textbf{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it's just confusing without a lot of additional explanation.}}

>>>>

\anna{In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait. However, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward. LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation. This procedure does not generalize beyond the linear regression case where the outcome is continuous. Because of this, distinct inferential algorithms are required to extend the LMM modeling methodology to nonlinear models such as logistic regression. \citet{mandt2017sparse} generalize the LMM modeling paradigm to binary phenotypes in the interest of classification. They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference. However, to the best of our knowledge, the joint problem of sparse estimation in the presence of dependent observations has been limited to the linear regression case \citep{seeger2011large, vattikuti2014applying, rakitsch2013lasso}.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods} \label{sec:methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
\subsection{\anna{lasso}}
We begin by briefly reviewing the key methodological components of the lasso. In the usual linear regression set-up where there is a continuous response $\by \in \real^n$, an $n \times p$ design matrix $\bX$, and a parameter vector $\bbeta \in \real^p$, the ordinary lasso minimizes the objective function

\begin{equation}
    Q(\bbeta | \lam) = \frac{1}{2n}|| \by -  \beta_0 - \bX \bbeta ||_2^2 + \lam ||\bbeta||_1,
\end{equation}
where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm. The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data. The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values. The tuning parameter, $\lambda$, controls the tradeoff between these two components. For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0. In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of ordinary least squares (OLS). The lasso assumes observations are independent. However, the $\ell_1$ penalty of the lasso can also be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.

@@

%------------------------------%
\subsection{\anna{Genomewide analysis case study}}
%------------------------------%

@@

\anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods. In a real data scenario, the true number of causal SNPs would not be known a priori. As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion. As of yet there is no clear consensus on the optimal way to select a penalty parameter for penalized LMMs.}

@@

\anna{\textbf{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it would just make things more confusing without a lot of additional explanation.}}
